Continuous gravitational waves in the lab: recovering audio signals with a table-top~optical microphone

### 2021-08-31 paper/responseRound2/AJPResponseLetter.tex

\item p.7, right column, Section V, para 2: Just to be clear, what have you changed from the apparatus in earlier sections (if anything), other than the photodiode described below? It reads like you have changed something.
\begin{itemize}
\item The only thing we have changed is replacing the webcam with the photodiode so that the apparatus can be used to capture higher frequency speech and music. In this section we use the interferometer to capture more complicated signals, so we draw comparisons to laser microphones used in other fields. We have added a sentence to this paragraph to clarify the apparatus changes. 
\end{itemize}

\item p.8, left column: This paragraph adds very little new information to the topic of Section V, and interrupts the flow of the paper. I recommend shortening it. 
\begin{itemize}
\item Agreed. This paragraph was added in response to reviewer 3's request for us to mention other methods of noise reduction from the gravitational-wave literature. This seemed like the most appropriate place for it to go, however we agree that it is not closely related to the work presented here. We have shortened this discussion to a single sentence and moved the literature review to the Further Reading section in the Supplementary Material. 
\end{itemize}

\item p.9, Figure 8: It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs. Here's an idea: we could attach a direct link to an audio clip for each row, perhaps a 10 second recitation of a familiar phrase. It would be available to online readers via s simple click. This would enhance the discussion significantly.
\begin{itemize}
\item This sounds a great idea! We have collated the four sounds files and attached them in the email. Please let us know if you need any further details from us. 
\end{itemize}

\item p.10, left column, para 2: This seems like an unlikely approach. Why not simply increase the sensitivity, as described in the following sentence, and rotate the apparatus relative to the source?
\begin{itemize}
\item We have clarified this in the text. The gain in having data from two interferometer is that directional information can be estimated based on the time delay between the signal arriving at each interferometer, in a similar way to how we are able to tell the direction of a sound with our ears. This is also how gravitational-wave detectors estimate the direction of compact binary mergers based on the time-delay of the signal reaching two or more detectors. 
\end{itemize}

### 2021-09-23 paper/ifo-introduction.tex
moving citations to after punctuation at editor request

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe~\cite{GW150914}.
Gravitational waves are a prediction of Albert Einstein's theory of General Relativity; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO), Advanced Virgo, GEO600, and KAGRA use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{TTExhibit:2021}.

% CW 
To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers~\cite{GWTC-2:2020,NSBH:2021}.
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
%Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
%Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref~\cite{ScoX1O2Viterbi:2019}).

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
% I don't think it is novel to use an optical device as a microphone, should we make this clearer? - james
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~
\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we observe a wandering frequency signal and analyze it using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the Supplementary Material). % deleted ``accompaning'', the material is already supplementary - james
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
Further reading and resources are provided in the Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.

>>>>

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe.\cite{GW150914}
Gravitational waves are a prediction of Albert Einstein's theory of General Relativity; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO), Advanced Virgo, GEO600, and KAGRA use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences.\cite{TTExhibit:2021}

% CW 
To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers.\cite{GWTC-2:2020,NSBH:2021}
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion.\cite{xraybinaries:1997}
Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref.~\cite{ScoX1O2Viterbi:2019}).

In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data.\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019} 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we observe a wandering frequency signal and analyze it using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the Supplementary Material).
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
Further reading and resources are provided in the Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.
